Computer-Assisted Analysis of the \\ Anderson-Hájek Ontological Controversy

### 2015-03-31 Papers/InProgress/HajekAndersonControversy/main.tex

\usepackage{amsmath}
\usepackage{xspace}
\usepackage{url}
\usepackage{commands}
\usepackage{xcolor}
\usepackage{comment}
\usepackage{modallogics}
\usepackage[backend=bibtex,sorting=ydnt,maxnames=999,maxcitenames=999,maxbibnames=999,natbib]{biblatex}
\addbibresource[label=bibs]{Bibliography.bib}

@@

For both \textbf{constant domain semantics} and \textbf{varying domain
semantics}, the following results hold for \emph{Anderson's
Emendation} (cf. Appendix \ref{apx:Anderson}): T1, C
and T3' can be quickly automated (in logics \K, \K and \KB,
respectively); the axioms A4 and A5 are proven redundant\footnote{
An axiom $A$ is redundant w.r.t. a set of axioms $S$ iff $A$ is
entailed by $S$.  } (the former in logic \KFourB and the latter
already in \K); a trivial countermodel (with two worlds and  two
individuals) for MC is generated by Nitpick (for all mentioned
logics); all axioms and definitions are shown to be mutually
consistent.

@@

The controversy over the superfluousness of A4 and A5 indicates a
trend to reduce the ontological argument to its bare essentials. In
this regard,  already
\citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates
that, by taking a notion of \emph{defective} as primitive and defining
the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become
derivable. These claims have been confirmed by the automated theorem
provers (in logic \KFourB). Within the same trend, the alternative
proposed by \citet{bjordal99} (cf. Appendix \ref{apx:Bjordal})
achieves a high level of minimality. He takes the property of being
God-like as a primitive and defines (B:D1) the positive properties as
those properties necessarily possessed by every God-like being. He
then briefly indicates (B:L1) that B:D1 is logically equivalent, under
modal logic \SFour, to the conjunction of D1', A2', A3' and A4'. This
has been confirmed in the computer-assisted formalization: A2' and A3'
can be quickly automatically derived in logic \K. A4' can be proved in
logic \KT (i.e. assuming reflexivity of the accessibility relation).
For constant domain semantics, proving D1 is possible in logic \KFour,
whereas for varying domain semantics, a counter model can be found
even in logic \SFive. Conversely, the proof that B:D1 is entailed by
D1', A2', A3' and A4' is possible already in logic \K. The provers
also show that theorem T3' follows from B:D1, B:D2, B:D3, A:A1' and
A5' already in logic \KB. Bjordal's last paragraph briefly mentions
Hájek's ideas about the superfluousness of A5' and claims that it is
possible, with (unclear) additional modifications of the definitions,
to eliminate A5' from his theory as well. Without any additional
modification, the automated reasoners show that A5' is actually not
superfluous. All these results, with the exception of the
aforementioned countermodel for D1', hold for both constant and
varying domain semantics.

@@

>>>>

\usepackage{amsmath}
\usepackage{xspace}
\usepackage{url}
\usepackage{commands}
\usepackage{xcolor}
\usepackage{comment}
\usepackage{modallogics}
\usepackage[backend=biber,sorting=ydnt,maxnames=999,maxcitenames=999,maxbibnames=999,natbib]{biblatex}
\addbibresource[label=bibs]{Bibliography.bib}

@@

For both \textbf{constant domain semantics} and \textbf{varying domain
semantics}, the following results hold for \emph{Anderson's
Emendation} (cf. Appendix \ref{apx:Anderson}): T1, C
and T3' can be quickly automated (in logics \K, \K and \KB,
respectively); the axioms A4 and A5' are proven redundant\footnote{
An axiom $A$ is redundant w.r.t. a set of axioms $S$ iff $A$ is
entailed by $S$.  } (the former in logic \KFourB and the latter
already in \K); a trivial countermodel (with two worlds and  two
individuals) for MC is generated by Nitpick (for all mentioned
logics); all axioms and definitions are shown to be mutually
consistent.

@@

The controversy over the superfluousness of A4 and A5 indicates a
trend to reduce the ontological argument to its bare essentials. In
this regard,  already
\citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates
that, by taking a notion of \emph{defective} as primitive and defining
the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become
derivable. These claims have been confirmed by the automated theorem
provers (in logic \KFourB). Within the same trend, the alternative
proposed by \citet{bjordal99} (cf. Appendix \ref{apx:Bjordal})
achieves a high level of minimality. He takes the property of being
God-like as a primitive and defines (B:D1) the positive properties as
those properties necessarily possessed by every God-like being. He
then briefly indicates (B:L1) that B:D1 is logically equivalent, under
modal logic \SFour, to the conjunction of D1', A2', A3' and A4'. This
has been confirmed in the computer-assisted formalization: A2' and A3'
can be quickly automatically derived in logic \K. A4' can be proved in
logic \KT (i.e. assuming reflexivity of the accessibility relation).
For constant domain semantics, proving D1' is possible in logic \KFour,
whereas for varying domain semantics, a counter model can be found
even in logic \SFive. Conversely, the proof that B:D1 is entailed by
D1', A2', A3' and A4' is possible already in logic \K. The provers
also show that theorem T3' follows from B:D1, B:D2, B:D3, A:A1' and
A5' already in logic \KB. Bjordal's last paragraph briefly mentions
Hájek's ideas about the superfluousness of A5' and claims that it is
possible, with (unclear) additional modifications of the definitions,
to eliminate A5' from his theory as well. Without any additional
modification, the automated reasoners show that A5' is actually not
superfluous. All these results, with the exception of the
aforementioned countermodel for D1', hold for both constant and
varying domain semantics.

@@

\section{Summary}

\newcommand{\R}{Red.}
\newcommand{\S}{Sup.}
\newcommand{\P}{Prov.}
\newcommand{\CS}{Countersat.}
\newcommand{\U}{--}

\begin{tabular}{lccc}
Proof               & A4 & H:A4 & A5 & A5' & H:A5 & T3 & T3'  &  MC \\
\hline
Scott (const)       & 
Scott (var)
Anderson (const)    &     &      
Anderson (var)      
\AOE
\AOEH
\AOEHH
\AOEHHH
\end{tabular}
